AUTOMATED TRANSCRIPTION OF DRUMSOUNDS

### 2017-08-16 Latex/Paper/main.tex
Continued writing text and made matrix plot prettier.

On first sight extracting the time point of a drum hit for a polyphonic, mixed piece of music sounds like a difficult task.
In fact though there are multiple mathematically simple algorithm to achieve that.
For this research we took an algorithm commonly used for these kind of problems: the non-negative matrix factorization (NMF).

In section \ref{sec:algo} there will be a detailed explanation of how NMF works.

@@

The idea to use NMF for pattern recognition was first proposed by Lee and Seung in 1999 \cite{lee1999}.
As described in their paper NMF is a versatile technique to find features in a data set.
It can for example be used to recognize parts of a face, find semantic features in text or certain events in a piece of music.

@@

NMF is in its essence a pattern matching algorithm.
It decomposes the magnitude spectrum of a song into two matrices.
The first one contains the magnitude spectra of the individual components we are interested in and the second one contains the time-varying gain of these components, which tells us when each of the components for the first matrix is active.
The sum of both matrices makes up an approximation of the original spectrum. \cite{virtanen2007}

@@

>>>>

On first sight extracting the time point of a drum hit for a polyphonic, mixed piece of music sounds like a difficult task.
In fact though there are multiple mathematically simple algorithm to achieve that.
For this research we took an algorithm commonly used for these kind of problems: the non-negative matrix factorization (NMF).
Section \ref{sec:algo} will show a detailed explanation of how the NMF works.

@@

The idea to use NMF for pattern recognition was first proposed by Lee and Seung in 1999 \cite{lee1999}.
As described in their paper NMF is a versatile technique to find features in a data set.
Lee and Seung proposed in their paper that it can be used to recognize parts in faces, find semantic features in texts or find certain musical events in a piece of music.
During the last decade NMF was researched and applied in all the fields they mentioned and many more.

In the field of music information retrieval NMF has also been used in a variety of applications.
Most notable are the fields of source separation and music transcription.
For source separation the goal is to separate a piece of music containing a mix of multiple sound sources into the their components.
Virtanen \cite{virtanen2007} has researched this topic among many others.

@@

As mentioned before we used a non-negative matrix factorization to extract the time points of drum events from a drum track.
Generally speaking the NMF factorizes a matrix (usually called $\mathbf{V}$) into two smaller matrices (usually called $\mathbf{W}$ and $\mathbf{H}$) with the constraint that every matrix element must be non-negative.
In turn the sum of $\mathbf{W}$ and $\mathbf{H}$ is an approximation of $\mathbf{V}$ (see figure \ref{fig:NMF}).

\begin{figure}[htb]

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{figures/NMF}}
%  \vspace{2.0cm}
  \medskip
\end{minipage}

\caption{Matrix $\mathbf{V}$ can be factorized into two matrices $\mathbf{W}$ and $\mathbf{H}$, so that their sum approximates $\mathbf{V}$. \scriptsize{\textcopyright WikiMedia CC BY-SA 3.0}}
\label{fig:NMF}

\end{figure}

In our case we use the magnitude spectrum of a song as the $\mathbf{V}$ matrix and decomposed it into two matrices.
The first one contains the magnitude spectra of the individual components we are interested in ($\mathbf{W}$) and the second one contains the time-varying gain of these components ($\mathbf{H}$), which tells us when each of the components for the first matrix is active.
For that reason the $\mathbf{W}$ matrix is usually called component matrix and $\mathbf{H}$ is usually called activation matrix.
Again the sum of both of these matrices makes up an approximation of the original spectrum \cite{smaragdis2003}.
A visual representation of that can be seen in figure \ref{fig:MatrixOverview} (d).

For our use case we were not interested in finding all the components and their activations.
We were rather interested in finding the corresponding activation of a specific component, like that of a base drum or a hi-hat.
This is also possible using NMF.
In order to do so, we specified $\mathbf{W}$ as the spectrum of a base drum and $\mathbf{V}$ continued to be the spectrum of the whole song.
The NMF then solves for the matrix $\mathbf{H}$ which tells at what point in time the specified base drum spectrum appears in the song.
In that setup the NMF essentially works as a pattern matching algorithm.

\begin{figure}[htb]

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{figures/MatrixOverview}}
%  \vspace{2.0cm}
  \medskip
\end{minipage}

\caption{(a) Magnitude spectrum of a drum track. (b) Magnitude spectrum of a base drum. (c) Activation matrix of the base drum. (d) Approximated spectrum. \scriptsize{\textcopyright Own work CC BY-SA 3.0}}
\label{fig:MatrixOverview}

\end{figure}

@@

- we only uses drum tracks, but it also works with mixed music. problem: ground truth
